### 2021-09-09 analysis/joinRatioBiomass.Rmd
sept 9

---
title: "Calculate residue factors"
author:
- Glen W Armstrong^[Department of Renewable Resources, University of Alberta]
date: "19/03/2021"
output:
  pdf_document:
    number_sections: yes
    latex_engine: xelatex
    keep_tex: yes
  word_document: default
  html_document:
    df_print: paged
bibliography: historicalResidueSupply.bib
---

```{r setup, include=FALSE, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo=FALSE, message=FALSE}
#setwd("g:/Shared drives/FES_Data_Sharing (Forestry)/Manuscripts/Historical Residue Supply/Analysis")
library(tidyverse)
library(ggthemes)
library(gridExtra)
library(viridis)
library(skimr)
library(scales)
library(DescTools)
library(zoo)
```

## Taper

calculates important ratios based on taper functions

```{r stuff}
ratios <- read.csv(file = '../data/ratios.csv')
biomass <- read.csv(file = '../data/bioCalcOutput.csv') %>% select(-Province,-Height)
ratiosBiomass <- ratios %>% right_join(biomass,by=c("Ecozone","Genus","Species","Variety","DBH"))
ratiosBiomass <- ratiosBiomass %>% add_column(residueBiomass=0) %>% add_column(residueFactor=0)
ratiosBiomass$residueBiomass <- ratiosBiomass$branches + ratiosBiomass$foliage +
  (ratiosBiomass$topBarkRatio * ratiosBiomass$stem_bark) + 
  (ratiosBiomass$topVolRatio * ratiosBiomass$stem_wood)
ratiosBiomass$residueFactor <- ratiosBiomass$residueBiomass / ratiosBiomass$merchVol / 1000
#ratios
#biomass
#ratiosBiomass
#ratiosBiomass %>% filter(Species == "TRE",DBH == 20)
ratiosBiomassBorealPlains <- ratiosBiomass %>% filter(Ecozone == 9, Species == "BAN" | Species == "CON" | Species ==  "GLA" | Species ==  "MAR" | Species == "TRE")
#ratiosBiomassBorealPlains

ggplot(ratiosBiomassBorealPlains, aes(x=DBH, y=residueFactor, colour=Species)) + geom_point(shape=21)

ggplot(ratiosBiomassBorealPlains, aes(x=merchVol, y=residueFactor, colour=Species)) + geom_point(shape=21)

```