### 2021-09-29 analysis/interp.Rmd

---
title: "interp_games"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo=FALSE, message=FALSE}
#setwd("/home/gwa/Work/historical-residue/Analysis")
setwd("/mnt/chromeos/GoogleDrive/SharedDrives/FES_Data_Sharing_Forestry/Manuscripts/histressupp/analysis")

library(tidyverse)
library(janitor)

biomass <- read.csv(file = '../data/bioCalcOutput.csv')

t <- biomass %>% select(-Tree_num,-Province) %>%
  filter(Ecozone == 4) %>%
  filter(Genus == "PICE") %>%
  filter(Species == "GLA")

x <- t$DBH
y_sw <- t$stem_wood
y_sb <- t$stem_bark
y_b <- t$branches
y_f <- t$foliage

dbh <- 25.3

approx(x,y_sw,xout=dbh,rule=2)
approx(x,y_sb,xout=dbh,rule=2)
approx(x,y_b,xout=dbh,rule=2)
approx(x,y_f,xout=dbh,rule=2)

species_prop <- read.csv(file = '../data/species_proportions_full.csv') %>%
  select(TRM,max_species)
tabyl(species_prop$max_species)
```