### 2021-10-05 analysis/residual_ratios.Rmd

---
title: "Residual Ratios"
author:

- Glen W Armstrong^[Department of Renewable Resources, University of Alberta]

date: "`r format(Sys.time(), '%Y-%m-%d')`"
#date: "2021-09-16"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document:
    number_sections: yes
    latex_engine: xelatex
    keep_tex: yes
bibliography: historicalResidueSupply.bib
---

```{r setup, include=FALSE, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo=FALSE, message=FALSE}
setwd("/mnt/chromeos/GoogleDrive/SharedDrives/FES_Data_Sharing_Forestry/Manuscripts/histressupp/analysis")
library(tidyverse)
```

## Residual ratios

calculates important ratios based on taper functions

```{r parameters}

residual_ratios <-
  read.csv(file = '../data/residual_ratios.csv', fileEncoding = 'UTF-8-BOM')
residual_ratios$prop_of_137 <- residual_ratios$merch_ib_vol / residual_ratios$merch_ib_vol137
residual_ratios
save(residual_ratios,file="../data/residualRatios.RData")

```