### 2021-10-17 analysis/manuscript.Rmd
writing at full speed

---
title: Spatial and temporal variability in economic availability of residual biomass from timber harvest in west-central Alberta, Canada (1990-2015)
author:
  - Glen W. Armstrong:
      institute:
        - rr
      correspondence: "yes"
      email: gwa@ualberta.ca
  - Jay A. Anderson:
      email: jaya@ualberta.ca
      institute:
        - rees
  - Alex W. McPhee:
      email: awmcphee@ualberta.ca
      institute: 
        - rees
institute:
  - rr : Department of Renewable Resources, University of Alberta
  - rees: Department of Resource Economics and Environmental Sociology, University of Alberta
date: "`r format(Sys.time(), '%Y-%m-%d')`"
#date: "2021-10-05"
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
output:
  pdf_document:
    includes:
      in_header: "preamble.tex"
    number_sections: yes
    latex_engine: xelatex
    keep_tex: yes
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
  html_document:
    number_sections: yes
    df_print: paged
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
  word_document:
     pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
#csl: canadian-journal-of-forest-research
csl: biomass-and-bioenergy
bibliography: historicalResidueSupply.bib
---
\raggedright
<!-- \fontfamily{cmr} -->
<!-- \fontsize{12}{22} -->
<!-- \fontseries{b} -->
<!-- \selectfont -->

```{r setup, include=FALSE, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo=FALSE, message=FALSE}
setwd("g:/Shared drives/FES_Data_Sharing_Forestry/Manuscripts/histressupp")
#setwd("/mnt/chromeos/GoogleDrive/SharedDrives/FES_Data_Sharing_Forestry/Manuscripts/histressupp/analysis")
library(tidyverse)
library(ggthemes)
library(gridExtra)
library(viridis)
library(skimr)
library(scales)
library(DescTools)
library(zoo)
library(ezknitr)
#ezknit(file="analysis/manuscript.Rmd", out_dir="output", fig_dir="myfigs")
```

# Introduction

1. **Focus on cost variability associated with transportation and forest  characteristics**
2. **Based on historical harvest locations (1990-2015)**

<!-- ## Some obvious citations -->

<!-- @Onyango2019 -->

<!-- @Carriquiry2011 -->

<!-- @Miles2009 -->

<!-- @Dessbesell2017 -->

<!-- @Dymond2010 -->

<!-- @Lambert2005 -->

<!-- @Lo2021 -->

<!-- @Malladi2018 -->

<!-- @Mansuy2017 -->

<!-- @Mansuy2020 -->

<!-- @Schroder2019 -->

<!-- @Shabani2013 -->

<!-- @Shooshtarian2018 -->

<!-- @Yemshanov2018 -->

<!-- @Zheng2021 -->

<!-- @Ackom2010 -->

<!-- @Acuna2019 -->

<!-- @Boudewyn2007 -->

<!-- @Peltola2011 -->

## Justification

**Do we talk about the weaknesses of BRIMS and how we used 
this to improve?**

```{r BRIMS, echo=FALSE, fig.cap="BRIMS residue", out.width = '100%'}
knitr::include_graphics("../img/BRIMS.png")
```

* renewable fuels better than fossil fuels
* forest harvest residue just piled and burned
* forest road networks in Alberta radiate from location
* terrain is difficult
* simple circle or doughnut models of limited utility

# Materials and methods

```{r constants, echo=FALSE, include=FALSE}
#hourly_truck_rate <- 175 + 225 + 40 #place holder.
# info fromhttps://www.rogerstrucking.ca/services/rates-and-equipment/62-tractor-truck-hourly-rates
# 175 $/hr for tractor, 225 for walking floor trailer, 40 for operator

#hourly_truck_rate <- 175 # adjusted from Roger's numbers to be more in line with Roeser
trailer_capacity <- 100 #m^3 from https://www.newtontrailers.com/trailer_hire/biomass_moving_floor_trailer/-/54/0
hourly_truck_rate <- 175 #$/hr
solid_volume_factor <- 0.4 #from fortsv5
adj_trailer_capacity <- trailer_capacity * solid_volume_factor
specific_gravity <- 0.38 #lodgepole pine
odt_trailer <- specific_gravity * adj_trailer_capacity
truck_rate <- hourly_truck_rate / odt_trailer

#truck_rate <- 6.43 # $/odt/hr
prepiling_cost <- 6.30 # $/odt
comminution_cost <- 25.60 # $/odt
other_cost <- 8.48 # $/odt
constant_cost <- prepiling_cost +  comminution_cost + other_cost # $·Mg^-1^ (DM)
location <- c('Whitecourt','Edson','Hinton')
capacity <- c(200000,400000,800000)

merch_vol_adj_factor <- 2.77 # based on a wild ass guess
#merch_vol_adj_factor <- 2.38 #based on West Fraser Hinton
odmt_adj_factor <-  0.10

#get prop of dec/con by twp
standing_time <- 20

```

## Harvested area

Canada's National Forest Inventory System (NFIS) is a product of the Canadian
Council of Forest Ministers (CCFM).  One of the data sets provided by the NFIS
identifies locations of forest harvesting across Canada for the years 1985-2015
detected from Landsat satellite imagery.  These data are at a 30 m spatial
resolution.  The data were downloaded from
https://opendata.nfis.org/downloads/forest_change/CA_forest_harvest_mask_year_1985_2015.zip
(accessed 2020-03-05). The data set is described in Hermosilla *et al.* (2016)
@Hermosilla2016.
The data were clipped to the Alberta provincial boundary which was obtained from
Altalis (https://www.altalis.com, accessed 2021-09-10), which manages much of
the spatial information belonging to the Alberta government.

## Alberta township coverage
For our analysis, most of our information was aggregated to the township level.
In Alberta, the legal description of any parcel of land uses the Alberta
Township Survey System (see
https://www.alberta.ca/alberta-township-survey-system.aspx, accessed
2021-10-08). Most townships are roughly square in shape, with an average area of
about 9,533 ha.  The GIS coverage for township outlines was obtained from
Altalis (https://www.altalis.com, accessed 2021-09-10).
The harvest areas obtained from the NFIS forest harvesting data set were
aggregated to the township level so that the harvest area for each township in
each of the years from 1985 to 2015 was recorded in a data set.

## Hypothetical plant locations and capacities
There are a number of companies with Forest Management Agreement Areas
located near Edson, Whitecourt, and Hinton.
The West Fraser Mills Ltd. (Hinton) FMA is the primary timbershed for the
Hinton Wood Products sawmill and the Hinton Pulp pulp mill.  The West Fraser
Mills Ltd (Edson) FMA is the primary timbershed for the Edson Forest Products 
sawmill.  The Weyerhaeuser Company Limited (Pembina Timberlands) FMA supplies 
the Edson oriented strand board mill and the Drayton Valley sawmill.  The 
Millar Western FMA supplies sawmills in Whitecourt and Fox Creek, a pulp mill
in Whitecourt, and a specialty wood products facility and Acheson.  The Blue
Ridge Lumber Inc. FMA supplies a sawmill in Blue Ridge, which is about 23 km 
from Whitecourt.  The ANC Timber Ltd. FMA supplies logs to several sawmills in
the area, in exchange for chips used to feed the Alberta Newsprint
Company newsprint mill in Whitecourt.

```{r west_central_ab, echo=FALSE, fig.cap="Location of Forest Management Agreement Areas serving Hinton, Edson, and Whitecourt.", out.width = '80%'}
knitr::include_graphics("../img/west_central_AB_2.png")
```

For this study, we assumed that forest harvest residue processing would take
place at a plant co-located with an existing forest products manufacturing
facility. We used the Weyerhaeuser oriented strand board plant in Edson, the
Millar Western Forest Products Ltd. pulp mill in Whitecourt, and the West Fraser
pulp mill in Hinton as the locations for the hypothetical forest residue
processing facilities.

As well as considering locations for the hypothetical plants, we considered 3
different plant capacities: 200 000 Mg/a (DM) corresponding to a small wood pellet
mill, 400 000 Mg/a (DM) corresponding to a large pellet mill or a small ethanol
plant, and 800 000 Mg/a (DM) corresponding to a large ethanol plant.

```{r bb, echo=FALSE, fig.cap="Boreal Bioenergy--Obed.", out.width = '80%'}
knitr::include_graphics("../img/BorealBioenergy.png")
```

## Administrative and ecological boundaries.

### FMU map

Many forest products companies operate in Alberta, and have different agreements
with the government with respect to how the forest is managed.  Some companies
have harvesting rights to just the needleleaf species, others have rights to
just the broadleaf species, and some have rights to both needleleaf and
broadleaf.

Utilization standards in Alberta are specified on the basis of stump height,
minimum stump diameter (outside bark), and minimum log length to a specified top
diameter (inside bark).  Utilization standards vary by FMU.  Stump heights used
in Alberta are 15 and 30 cm; stump diameters are 13 or 15 cm; top diameters are
7, 10, 11, or 12 cm; and minimum log lengths are 2.44, 2.66, 2.67, 3.55, 3.76, 
or 4.88 m.  These utilization standards are necessary to relate volumes 
calculated to the 13/7 standard for the National Forest Inventory to the harvest
volumes reported by the companies.  Perhaps most importantly, the minimum top
diameter will have an effect on the volume of residue left in the forest 
corresponding to the tops of trees.

```{r FMU, echo=FALSE, fig.cap="Location of Forest Management Units in Alberta showing Green and White Areas.", out.width = '80%'}
knitr::include_graphics("../img/FMU.png")
```

### Terrestrial ecozone map

The NFIS' individual tree biomass calculator @NFIS2016 uses province, ecozone, 
species, dbh, and height as input variables, so ecozone is important.  It is based on 
equations developed by Lambert
*et al.* (2005) @Lambert2005.

```{r ecozones, echo=FALSE, fig.cap="Terrestrial ecozones of Canada within Alberta.", out.width = '80%'}
knitr::include_graphics("../img/ecozones.png")
```
Example output from the biomass calculator.

```{r biomass_calculator_example, echo=FALSE, fig.cap="Biomass calculator example.", out.width = '80%'}
knitr::include_graphics("../img/biomass_calc_example.png")
```

```{r taper, echo=FALSE, fig.cap="Volume calcualation example.", out.width = '80%'}
knitr::include_graphics("../img/taper.png")
```
## Spatial variability in residual biomass production

There is considerable spatial variability in the amount of timber harvest
residues available for biomass collection.  The variability relates to the
distribution of harvested species and age classes across the landscape, to the
timber harvesting practices of different operators, and to other factors.
We used the 250 m resolution raster maps of Canada's forest attributes for 2011
from Natural Resources Canada @Beaudoin2017 to retrieve information on species
composition, age classes, and merchantable volume.  The data were clipped to the
Alberta boundary.  Merchantable volumes for Alberta in this dataset were
compiled to a 13/7 utilization standard, meaning that stump height was assumed
to be 30 cm, that stump diameter (outside bark) needed to be at least 13 cm, and
that the length of the log to a 7 cm inside bark diameter must be at least 4.88
m for a log to be merchantable.  This is one of the utilization standards used
in Alberta and was used for Canada's National Forest Inventory (NFI henceforth)
@Boudewyn2007.

The first step was to identify cells that contained merchantable forest
inventory. We selected cells that contained at least 50 m^3^/ha of merchantable
volume, as that corresponds with merchantability limits specified in Alberta
@Alberta2016. We filtered out cells with more than 50% crown closure of larch
(*Larix spp.* Mill.) and birch (*Betula spp.* L.) as those genera are rarely harvested at
a commercial scale in Alberta.  We filtered out cells with stand age less than
80 years as it is unusual to harvest stands younger than that in Alberta.
This filtered raster was "unioned" with the township coverage, and an average of
the merchantable volume in the merchantable cells was calculated.  This volume
was adjusted upwards by the ratio of the average volume per hectare harvested
from the West Fraser (Hinton) Forest Management Agreement Area (FMA) to the
average merchantable volume for each township calculated from the filtered
raster and township coverages.  The multiplicative volume adjustment factor was
calculated as `r sprintf("%5.2f",merch_vol_adj_factor)`.

**A little more explanation is needed..**

Something about needleleaf and broadleaf trees.  Most common species harvested
are white spruce (*Picea glauca* (Moench) Voss), lodgepole pine 
(*Pinus contorta* var. *latifolia* Engelm.), jack pine 
(*Pinus banksiana* Lamb.), trembling aspen (*Populus tremuloides* Michx.),
and less commonly black spruce 
(*Picea mariana* (Mill.) Britton, Sterns & Poggenburg), Engelmann spruce 
(*Picea engelmannii* 
Parry ex Engelm.),  balsam poplar
(*Populus balsamifera* L.), 
balsam fir (*Abies balsamea* (L.) Mill.),  
subalpine fir (*Abies lasiocarpa* (Hooker) Nuttall), 
and Rocky Mountain Douglas-fir 
(*Pseudotsuga menziesii* var. *glauca*  (Mayr) Franco).

## Service areas

The provincial road network was sorted into four classes: primary highways,
other paved roads, two-lane gravel roads, and one-lane gravel or dirt roads.  
Loaded and unloaded travel speeds were assumed for each road class, according
to the table below. The harmonic mean of loaded and unloaded speeds 
(rounded to the nearest
multiple of 5 km/h) were calculated as

$s_h= \left(\frac{s_l^{-1} + s_u^{-1}}{2}\right)^{-1}$.

Use of harmonic mean allows us to calculate the correct round trip travel time.

| Road class      | loaded | unloaded | harmonic mean | 
|:----------------|-------:|---------:|--------------:|
| primary highways|    80  |       95 |           85  |
| other paved     |    65  |       80 |           70  |
| 2-lane gravel   |    50  |       70 |           60  | 
| 1-lane gravel or dirt  |    20  |       30 |           25  | 
| no road           |     2  |       2  |           2   |

Check out https://haozhu233.github.io/kableExtra/awesome_table_in_html.html to
try to format this table better.

* Rasterized network areas were generated using the QGIS plugin, QNEAT3
(Iso-Area as Interpolation) @Raffler2018. **How big are the cells?**

* QGIS zonal statistics were used to calculate the median network travel time in
each township.

```{r haul, echo=FALSE, fig.cap="Alberta road network (left) and one-way haul time from Edson, limited to 6 hours.", out.width = '100%'}
knitr::include_graphics("../img/network_service.png")
```

# Theory/Calculation

This is a second best analysis.  Detailed information on harvest volumes
tied to spatial location is held by the provincial government as scale records.
Real (as opposed to modelled) forest inventory information is collected by
holders of forest management agreements as Alberta Vegetation Inventory.
Despite the fact that forest harvesting occurs mostly on public land, the
government considers this information to be the property of companies, and 
requires the consent of companies to release it to researchers.  There are
currently 21 Forest Management Agreements in Alberta.  Permission would need 
to be obtained from the holders of each of those FMAs as well as smaller
companies that operate sawmills.

The GIS analysis for this paper was done using QGIS @QGISorg2021.  Data
summaries and statistical analysis were done using R @R2021 with RStudio
@RStudioTeam2021, and tidyverse @Wickham2019.

## Residue yield (Mg · ha^-1^ (DM)) by township.

### Merchantable volume township

1. Filter out unmerchantable cells from 250 m data
   1. age < 80 years
   2. merchantable volume < 50 m^3^/ha based on 13+/7 cm utilization standard
   8. Bw + Lt prop > 50
2. average merchantable volume of remaining cells to the township level
3. adjust average inventory volume to actual harvested volume from the 
national forestry database @CCFM2021.
   - average volume/ha harvested / inventory volume = 2.77

### Tree characteristics

Most of he forest management plans prepared by FMA holders include projections
of average harvested pieces size, for both needleleafed and broadleafed trees,
if relevant.  We used the piece size reported for the
first 5-year period of the 200-year planning horizon.

We used the harvest piece sizes reported in the detailed forest management
plans for the townships in each FMAA.  For those areas where piece size was
not reported we assumed an average piece size of 0.278 m^3^ for needleleafed
trees and 0.264 m^3^ for broadleafed trees.  These are the average of the values
reported in the forest management plans.  These values are based on the 
utilization
standards used on the FMA.

Utilization standards vary by FMA, and are specified as combinations of 
stump height, stump diameter (outside bark), and minimum log length to a 
specified top diameter.  Stump heights used include 15 cm and 30 cm; stump 
diameters include 15 and 13 cm; top diameters include 12, 11, 10, and 7 cm;
minimum log lengths include 2.44, 2.67, 3.66, 3.76, and 4.88 m.  For areas 
outside of FMAAs we use 30 cm stump height, 15 cm minumum stump diameter, 11
cm top for needleleafed trees and 10 cm top for broadleafed trees, and a 4.88
m minimum log length.  These correspond to the default utilization standards
specified in the Alberta timber harvest planning and operating ground rules
@Alberta2016.

Every township has a leading needleleafed and broadleafed species and
a proportion of merchantable volume associated 
with it derived from the 250 m resolution national inventory @Beaudoin2017.

Ung *et al.* @Ung2013 present taper models and the associated coefficients
for the tree species found in Canada.  These taper models can be used to
find the diameter at any point along the main bole of a tree given species
and diameter at breast height (DBH).  DBH is a commonly used tree measurement
and is the diameter of the tree measured at 1.3 m above ground level, in
most jurisdictions.  It is straightforward to determine cross-sectional area
of the tree at any height.  Integrating this area function will yield volume
between any two heights.

\[
d(d_b,h) = \left(d^2_{b} \frac{\beta_0 d_{b}^{\beta_1}}{\beta_0 d_{b}^{\beta_1} - 1.3} \left(\frac{h}{1.3}\right)^{2-\beta_2}\right)^{-2}
\]

where $d$ is the diameter (cm) at height $h$ (m), and $d_b$ is the diameter at
breast height (1.3 m above ground level)

Implicit in this equation is an equation for tree height

\[
H(d_b) = \beta_0 d_b ^ {\beta_1}
\]

The cross-sectional area equation

\[
a(d_b,h) = \pi \left(\frac{d(d_b,h)}{200}\right)^2
\]

\[
v(d_b,h_s,h_t) = \int_{h_s}^{h_t}\pi \left(\frac{d(d_b,h)}{200}\right)^2dh
\]

The constant 200 is used to convert a diameter (cm) to a radius (m).  
**find a way to reduce possible confusion with $d$**

There is also a simple linear model relating outside bark to inside bark diameter.  The
taper function is for outside bark diameter, but piece sizes are reported
as inside bark volume.  Binary search is used to find the height associated
with the inside bark diameter specified in the utilization standard.

Every combination of township and FMU has a piece size for needleleafed and
broadleafed associated with it, and the leading needleleafed and broadleafed 
species.  We used a binary search procedure to find dbh and height of the 
"average" broadleafed and needleleafed tree in each township based on the 
companies utilization standard.  An average piece size to the 13+/7 utilization
standard was also calculated by integrating the area function using the 
appropriate limits.

Using these equations, we are able to calculate wood volumes and bark volumes
between any two heights along the tree stem.  We calculate the proportions 
of total tree and bark volumes in the top of the tree.  We use this volume 
proportion later to calculate biomass of stemwood and bark in the top.

The number of harvested stems per ha is calculated by dividing the 13/7 volume
per ha by the 13/7 piece size.

### Individual tree biomass.

Canada's national forest inventory website provides a useful individual
tree biomass calculator @NFIS2016.  It calculates the biomass of tree 
components (stem wood, stem bark, branches, foliage, and total) using
province/territory, terrestrial ecozon, species, dbh, and tree height as input.
It is possible to upload a file to the calculator so that the biomass of
many trees can be calculated at once.  The calculator is based on work
by Lambert *et al.* @Lambert2005.

We assume that the residual biomass is in the stem wood and stem bark from the
tops of the harvested trees and the branches and foliage.  The residual biomass
from the tops is calculated using the total biomass of stem wood and stem bark
adjusted by the volume proportions calculated above.

### Township level biomass

For both needleleaf and broadleaf, we multiply the residue associated with the
individual tree by the number of harvested stems per ha.  We reduce this
value using a recovery factor of 0.645 **??** following Peltola *et al.* 
@Peltola2011, recognizing that some of the residual biomass will 
be unrecoverable.

```{r resdiue, echo=FALSE, fig.cap="Biomass residue available from harvest.", out.width = '100%'}
knitr::include_graphics("../img/residue_per_ha.png")
```

## Cost calculations

```{r roser, echo=FALSE, fig.cap="Röser.", out.width = '100%'}
knitr::include_graphics("../img/roser.png")
```

```{r hourly_rates, echo=FALSE, fig.cap="Hourly rates for forestry equipment.", out.width = '100%'}
knitr::include_graphics("../img/hourly_rates.png")
```

Following @Roser2013, the pre-piling cost was set to
`r sprintf("%5.2f",prepiling_cost)` \$·Mg^-1^ (DM), the comminution (*grinding ?*) cost
was set to `r sprintf("%5.2f",comminution_cost)`  \$·Mg^-1^ (DM), and other costs were
set to  `r sprintf("%5.2f",other_cost)` \$·Mg^-1^ (DM).  The total "constant" costs are
`r sprintf("%5.2f",constant_cost)` \$·Mg^-1^ (DM).
Using *blah blah blah* , the truck rate was set to 
`r sprintf("%5.2f",truck_rate)` \$·Mg^-1^ (DM)/h.
A 100 m³ trailer can haul `r odt_trailer` Mg (DM) of chipped residue.

```{r read_data_files, echo=FALSE}
newAB <- read.csv(file = '../data/AlbertaHarvestLevels.csv',
                  fileEncoding = 'UTF-8-BOM')
t <- read.csv(file = '../data/township_merchantable_volume.csv')
twp_merch_vol <- t %>% select(c(TRM,mean_merch_volume)) 
twp_merch_vol$adj_vol <- twp_merch_vol$mean_merch_volume * merch_vol_adj_factor
#twp_merch_vol$residue_odmt <- twp_merch_vol$adj_vol * odmt_adj_factor
load("../data/biomass_by_trm.RData")
twp_merch_vol_temp <- merge(twp_merch_vol,biomass_by_trm,by = "TRM", all = TRUE)
twp_merch_vol <- twp_merch_vol_temp %>% 
  mutate(adj_factor = res_total/adj_vol)

mean(twp_merch_vol$adj_factor)

t <- read.csv(file = '../data/harvest_township_stats_melt_area.csv')
t <- subset(t, select = c(TRM,year,harvest_ha)) %>% filter(year >=1990)
t <- merge(t, twp_merch_vol, by = "TRM", all=TRUE)
t$adj_vol <- t$adj_vol %>% replace_na(0)  # there is an NAs problem here.  replace NAs with average.
                                              # is that the right way to do it?
t$unadj_harvest_vol <-t$harvest_ha * t$mean_merch_volume
t$harvest_vol <- t$harvest_ha * t$adj_vol 
t$residue_odmt <- t$harvest_ha * t$res_total
t <- t %>% 
  filter(!is.na(residue_odmt))

harv <- subset(t, harvest_ha > 0)
#head(harv)
#summary(harv$harvest_ha)
#summary(garv$harvest_vol) 
#summary(harv$adj_vol)

#harv2 <- aggregate(harv$harvest_vol,by=list(Category=harv$year), FUN=sum)
harv2 <- harv %>% group_by(year) %>%
  summarise(harvest_vol = sum(harvest_vol),
            unadj_harvest_vol = sum(unadj_harvest_vol))

# %>%
#   mutate(hvol2=rollapply(harvest_vol,5,mean,align='left',fill=NA))

#summary(harv2$harvest_vol)
#harv2
# ggplot(data=harv2, aes(x=year)) +
#   geom_line(aes(y = harvest_vol), colour="darkred") +
# #  geom_line(aes(y = harvvol), colour="steelblue") +
#   labs(x = "year") + labs(y = "harvest volume (million m3)")  +
#   scale_y_continuous(limits=c(0,30e6),labels = label_number(scale = 1/1e6))

#2009 is the median harvest volume year for Alberta in 1990-2015

# ggplot(data=harv2, aes(x=year,y=harvest_vol)) +
#   labs(x = "year") +
#   labs(y = "harvest volume (million m3)") +
#   geom_line()  +
#   scale_y_continuous(limits=c(0,30e6),labels = label_number(scale = 1/1e6))
# +
#   scale_colour_viridis(discrete = TRUE) +
#   theme(legend.position = "none") +
#   facet_grid(cols=vars(location)) +
#   scale_x_continuous(limits=c(0,800000),labels = label_number(scale = 1/1000))
harv3 <- read.csv(file = '../data/ABHarvestVol.csv') %>% filter(year >=1990)
#harv3

harv4 <- merge(harv2, harv3, by = "year", all=TRUE)
harv5 <- merge(harv4, newAB, by = "year", all=TRUE)

harv_comparison <- ggplot(data=harv5, aes(x=year,y=harvest_vol, colour="Estimated harvest")) +
  labs(x = "year") +
  labs(y = bquote('harvest volume' ~(10^6 ~m^3))) +
  geom_line()  +
  geom_line(aes(y=total, colour="Actual harvest"))  +
  scale_y_continuous(limits=c(0,30e6),labels = label_number(scale = 1/1e6))
# +
#   scale_colour_viridis(discrete = TRUE) +
#   theme(legend.position = "none") +
#   facet_grid(cols=vars(location)) +
#   scale_x_continuous(limits=c(0,800000),labels = label_number(scale = 1/1000))
# 
# head(harv4)

t <- read.csv(file = '../data/interpolated_median_oneway_times_seconds.csv')

d <- pivot_longer(t,cols=c("edson_median","hinton_median","whitecourt_median"),names_to = "location")
d <- d %>% mutate(location=recode(location,whitecourt_median="Whitecourt",edson_median="Edson",hinton_median="Hinton"))
#d <- select(d, -sum)
d$cycleTime <- 2* (d$value/3600) + standing_time/60

d$residueCost <- d$cycleTime * truck_rate + prepiling_cost + comminution_cost + other_cost
d <- select(d, -value)

#head(d)

big_joined <- inner_join(d,harv, by = "TRM")
#head(big_joined)
#tail(big_joined)

cum.odmt <- big_joined %>%
  arrange(location,year,cycleTime) %>%
  group_by(location,year) %>%
  mutate(residue.cumsum=cumsum(residue_odmt))

```

As a check, we obtained harvest volumes from Alberta for the period 1990-2015 
from the national
forestry database @CCFM2021.  
Note that in 1998, industrial salvage volumes of 369 791 m^3^ of
softwood and 264 647 m^3^ are not included in the harvest levels 
shown in the graph.

```{r pressure, echo=FALSE, fig.cap="NFDP harvest levels compared with estimated.", out.width = '100%'}
#knitr::include_graphics("../img/alberta_harvest.png")
harv_comparison
```

```{r harvbyowner, echo=FALSE, fig.cap="NFDP harvest levels by ownership.", out.width = '100%'}
knitr::include_graphics("../img/harvestbyowner.png")
```

Some lovely supply curves

```{r supply_curves, echo=FALSE, warning=FALSE, fig.cap="Residue supply (cycle time) curves by location.  Each curve represents one year.  The red curve is for 2009, which was the year of median harvest volume."}

p <- ggplot(data=cum.odmt, aes(x=residue.cumsum,y=cycleTime,group=year,colour=as.factor(year))) +
  labs(x = "cumulative residue (1000 Mg (DM))") +
  labs(y = "cycle time (hr)") +
  geom_line() +
  scale_colour_manual(values=c("Grey","Grey","Grey","Grey","Grey","Grey","Grey",
                               "Grey","Grey","Grey","Grey","Grey","Grey","Grey",
                               "Grey","Grey","Grey","Grey","Grey","Red","Grey",
                               "Grey","Grey","Grey","Grey","Grey","Red")) +
#  scale_colour_viridis(discrete = TRUE) +
  geom_line(data=cum.odmt[cum.odmt$year==2009,],aes(x=residue.cumsum,y=cycleTime, colour="Red"))+
  theme(legend.position = "none") +
  facet_grid(cols=vars(location)) +
  scale_x_continuous(limits=c(0,800000),labels = label_number(scale = 1/1000))+
  scale_y_continuous(limits=c(0,12))

p

```

```{r more_supply_curves, echo=FALSE, warning=FALSE, fig.cap="Residue supply (delivered cost) curves by location.  Each curve represents one year.  The red curve is for 2009, which was the year of median harvest volume."}
ggplot(data=cum.odmt, aes(x=residue.cumsum,y=residueCost,group=year,colour=as.factor(year))) +
  labs(x = "cumulative residue (1000 Mg (DM))") +
  labs(y = "marginal cost ($·Mg^-1^ (DM))") +
  geom_line() +
  scale_colour_manual(values=c("Grey","Grey","Grey","Grey","Grey","Grey","Grey",
                               "Grey","Grey","Grey","Grey","Grey","Grey","Grey",
                               "Grey","Grey","Grey","Grey","Grey","Red","Grey",
                               "Grey","Grey","Grey","Grey","Grey","Red")) +
#  scale_colour_viridis(discrete = TRUE) +
  geom_line(data=cum.odmt[cum.odmt$year==2009,],aes(x=residue.cumsum,y=residueCost, colour="Red"))+
  theme(legend.position = "none") +
  facet_grid(cols=vars(location)) +
  scale_x_continuous(limits=c(0,800000),labels = label_number(scale = 1/1000)) +
  scale_y_continuous(limits=c(50,175)) 
```

```{r boxplots, echo=FALSE, warning=FALSE, fig.cap="Distribution of residue costs by location, capacity, and year"}
t200 <- cum.odmt %>% filter(residue.cumsum <= 200000)
t200$capacity <- 200
t400 <- cum.odmt %>% filter(residue.cumsum <= 400000)
t400$capacity <- 400
t800 <- cum.odmt %>% filter(residue.cumsum <= 800000)
t800$capacity <- 800

t <- bind_rows(t200,t400)
t <- bind_rows(t,t800)

t <- t %>%
  arrange(year,location, cycleTime)

p <- ggplot(t,aes(factor(year),residueCost))  + 
  geom_tufteboxplot() +  
  facet_grid(cols=vars(location),rows=vars(capacity)) +
  labs(y = "residue cost ($·Mg^-1^ (DM))", x="year") +
  scale_x_discrete(breaks=c("1990","","","","","1995","","","","","2000","","","","","2005","","","","","2010","","","","","2015")) +
  theme(axis.text.x = element_text(angle=90))
p

```

```{r avecosts, echo=FALSE, warning=FALSE, fig.cap="Average costs by location, capacity, and year "}

t <- t %>%
  arrange(location, capacity, year)

avet <-t %>%
  group_by(location,capacity, year) %>%
  summarize(avecost = mean(residueCost))

ggplot(data=avet, aes(x=year,y=avecost,colour=location))+
  labs(x = "year") +
  labs(y = "average cost ($·Mg^-1^ (DM))") +
  geom_line() +
  facet_grid(rows=vars(capacity)) +
  scale_y_continuous(limits=c(60,150))

```

# Results and discussion

Whitecourt is low cost for most years at all capacities.  Noticable spike in
1999.  Probably due to Virginia Hills fire salvage.  Note that this spike
also affects Hinton and Edson at 800 000 Mg/a (DM) capacity, 
because they reach into the 
Whitecourt area.

As expected, residue costs are lower with lower capacities because they don't have to reach
so far out.

# Conclusions

# Data availability

# Acknowledgements

This research was supported by funding from the Canada First Research Excellence
Fund as part of the University of Alberta’s Future Energy Systems 
research initiative
<!-- (Project # CFREF-2015-00001) -->
.

Dennis Gray contributed to data compilation.

We work with a group of talented people.  Thanks to Marty Luckert, Feng Qiu,
Grant Hauer, and Wenbei Zhang for helpful discussion.

# Reference list