### 2021-11-23 analysis/newfigs.Rmd

---
title: Spatial and temporal variability in economic availability of residual biomass
  from timber harvest in west-central Alberta, Canada (1990--2015)
subtitle: 'Tables and Figures'
author:
- Glen W. Armstrong:
    institute: rr
    correspondence: 'yes'
    email: gwa@ualberta.ca
- Jay A. Anderson:
    email: jaya@ualberta.ca
    institute: rees
- Alex W. McPhee:
    email: awmcphee@ualberta.ca
    institute: rees
institute:
- rr: Department of Renewable Resources, University of Alberta, Edmonton AB  T6G 2H1,
    Canada
- rees: Department of Resource Economics and Environmental Sociology, University of
    Alberta, Edmonton AB  T6G 2H1, Canada
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  pdf_document:
    includes:
      in_header: preamble.tex 
    number_sections: yes
    latex_engine: xelatex
    keep_tex: yes
    pandoc_args:
    - --lua-filter=scholarly-metadata.lua
    - --lua-filter=author-info-blocks.lua
  bookdown::pdf_document2: default
  html_document:
    number_sections: yes
    df_print: paged
    pandoc_args:
    - --lua-filter=scholarly-metadata.lua
    - --lua-filter=author-info-blocks.lua
  word_document:
    number_sections: yes
    pandoc_args:
    - --lua-filter=scholarly-metadata.lua
    - --lua-filter=author-info-blocks.lua
  bookdown::html_document2: default
header-includes:
- \usepackage[left]{lineno}
- \usepackage{setspace}
- \linenumbers
geometry: left=1.5in,right=0.5in,top=1in,bottom=1in
csl: biomass-and-bioenergy
bibliography: historicalResidueSupply.bib
---

\raggedright

```{r setup, include=FALSE, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo=FALSE, message=FALSE}
#setwd("G:/Shared drives/FES_Data_Sharing_Forestry/Manuscripts/histressupp/analysis")
#setwd("/mnt/g/Shared drives/FES_Data_Sharing_Forestry/Manuscripts/histressupp/analysis")
#setwd("/mnt/chromeos/GoogleDrive/SharedDrives/FES_Data_Sharing_Forestry/Manuscripts/histressupp/analysis")
library(tidyverse)
library(ggthemes)
library(scales)
library(gt)
library(quantmod)
library(bookdown)
```

\newpage

# Tables {-}

Table 1:  Road speeds based on FPInnovations (km/h).

| Road class      | loaded | unloaded | harmonic mean | 
|:----------------|-------:|---------:|--------------:|
| primary highways|    85  |       95 |           90  |
| other paved     |    77  |       86 |           81  |
| 2-lane gravel   |    60  |       67 |           63  | 
| 1-lane gravel or dirt  |    29  |       25 |           27  | 
| no road           |     5  |       5  |           5   |

Table 2:  Summary statistics.

<!-- ```{r ave, echo=FALSE} -->
<!-- ave -->
<!-- ```  -->

\newpage

# Figures {-}

```{r canada, echo=FALSE, fig.cap="Location of Alberta in Canada.", out.width = '100%'}
knitr::include_graphics("../img/canada.png")
```

```{r cutblocks2009, echo=FALSE, fig.cap="Harvest areas for 2009.", out.width = '80%'}
knitr::include_graphics("../img/cutblocks2009.png")
```

```{r harvesttwp, echo=FALSE, fig.cap="Harvest areas for 2009 with aggregated to township level.  Only those townships where the proportion of total area harvested was 0.001 or greater are shown", out.width = '80%'}
knitr::include_graphics("../img/harvesttwp.png")
```

\newpage
\blandscape
```{r merchvol, echo=FALSE, fig.cap="Merchantable volume by township.  The left panel indicates total merchantable volume in the township from the 250 m data, and the boundaries of Alberta's green and white areas.", out.width = '80%'}
knitr::include_graphics("../img/Merchantable Volume.png")
```
\elandscape

\newpage
\blandscape
```{r haul, echo=FALSE, fig.cap="Alberta road network (left) and cycle time (right) from Whitecourt, limited to 13 hours.", out.width = '100%'}
knitr::include_graphics("../img/network_service.png")
```
\elandscape

<!-- ```{r studyarea, echo=FALSE, fig.cap="Location of hypothetical residue processing plants in relation to FMAA boundaries and the Green and White Areas of Alberta.", out.width = '80%'} -->
<!-- knitr::include_graphics("../img/study_area.png") -->
<!-- ``` -->

<!-- ```{r ecozones, echo=FALSE, fig.cap="Terrestrial ecozones of Canada within Alberta.", out.width = '80%'} -->
<!-- knitr::include_graphics("../img/ecozonesAlex.png") -->
<!-- ``` -->

\newpage
\blandscape
```{r pressure, echo=FALSE, fig.cap="Actual harvest levels from the national forestry database compared with estimated values.", out.width = '100%'}
knitr::include_graphics("../img/harv_comparison.png")
#suppressWarnings(print(harv_comparison))
```
\elandscape

<!-- ```{r residue, echo=FALSE, fig.cap="Biomass residue available from harvest.", out.width = '100%'} -->
<!-- knitr::include_graphics("../img/residue.png") -->
<!-- ``` -->

\newpage
\blandscape
```{r cyclePlot, echo=FALSE, fig.cap="Distribution of cycle times by location for each of the years 1990--2015.  Black curves are for a low harvest level year (1990), a medium harvest level year (2009), and a high harvest level year (2005).  Gray curves are for each of the other 23 years.", out.width = '100%'}
knitr::include_graphics("../img/cyclePlot.png")
#suppressWarnings(print(cyclePlot))
```
\elandscape

\newpage
\blandscape
```{r mcpPlot, echo=FALSE, fig.cap="Distribution of marginal costs by location for each of the years 1990--2015.  Black curves are for a low harvest level year (1990), a medium harvest level year (2009), and a high harvest level year (2005).  Gray curves are for each of the other 23 years.", out.width = '100%'}
knitr::include_graphics("../img/mcPlot.png")
#suppressWarnings(print(mcPlot))
```
\elandscape

\newpage
\blandscape
```{r bPlot, echo=FALSE, fig.cap="Boxplots of costs by location and biomass requirements for the years 1990--2015.", out.width = '100%'}
knitr::include_graphics("../img/bPlot.png")
#suppressWarnings(print(bPlot))
```
\elandscape

\newpage
\blandscape
```{r avePlot, echo=FALSE, fig.cap="Average costs by location and year.", out.width = '100%'}
knitr::include_graphics("../img/avePlot.png")
#suppressWarnings(print(avePlot))
```
\elandscape

```{r w2009, echo=FALSE, fig.cap="Relationship between timber harvest intensity and required supply area for facilities with biomass requirements of 200, 400, and 800 Gg/a.  The area of the dots reflects the proportion of the area of each township harvest in 2009", out.width = '100%'}
knitr::include_graphics("../img/w2009.png")
```

\newpage
\blandscape
```{r wcontour, echo=FALSE, fig.cap="Supply area for facilities in Whitecourt with biomass requirements of 200, 400, and 800 Gg/a for low cost (2005), medium cost (2009), and high cost (1990) years.", out.width = '100%'}
knitr::include_graphics("../img/Cumulative Contours Whitecourt.png")
```
\elandscape